---
title: "Flux Footprint Analysis with MATLAB"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flux Footprint Analysis with MATLAB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = TRUE
)
```

# Introduction

This vignette demonstrates flux footprint calculation using MATLAB's faster implementation of the Kljun et al. (2015) model, integrated with R for data preparation and visualization.

## Prerequisites

**Required:**
- MATLAB installed on your system
- MATLAB FFP functions from [http://footprint.kljun.net/](http://footprint.kljun.net/)
- R package `matlabr` installed: `install.packages("matlabr")`

## Setup
```{r setup}
library(FFPUtilities)
library(ggplot2)
library(sf)
library(lubridate)
library(matlabr)

# Define date range
start_date <- "2020-05-15"
end_date <- "2020-05-16"

# Get MATLAB scripts directory from package
matlab_ffp_path <- system.file("FFP_Matlab", package = "FFPUtilities")

# Verify MATLAB files exist
if (matlab_ffp_path == "" || !file.exists(file.path(matlab_ffp_path, "calc_footprint_matlab.m"))) {
  stop("MATLAB scripts not found in package. ",
       "Please ensure FFP_Matlab folder is in inst/ directory.")
}

message("MATLAB scripts found at: ", matlab_ffp_path)

# Output directory for MATLAB results
output_dir <- "C:/Users/kono0001/OneDrive - Sveriges lantbruksuniversitet/OTHER/personal/helping_others/Betty/Skogaryd/Scripts/Temp"
```

# Step 1: Load and Prepare Data in R
```{r load-data}
# Load EC data
ec_file <- system.file("extdata", "ec_data.csv", package = "FFPUtilities")
ec_data <- read.csv(ec_file)
ec_data$datetime <- parse_date_time(
  ec_data$datetime, 
  orders = c("ymd HMS", "ymd"), 
  tz = "UTC"
)

# Load site coordinates
site_file <- system.file("extdata", "site_coords.csv", package = "FFPUtilities")
site_coords <- read.csv(site_file)

head(ec_data)
site_coords
```

# Step 2: Extract BLH from GRIB
```{r extract-blh}
grib_file <- system.file("extdata", "degfert_blh.grib", package = "FFPUtilities")

blh_data <- extract_blh_from_grib(
  grib_file = grib_file,
  site_long = site_coords$long,
  site_lat = site_coords$lat,
  filter_start_date = start_date,
  filter_end_date = end_date
)

head(blh_data)
```

# Step 3: Prepare Data for MATLAB
```{r prepare-data}
ec_clean <- prepare_EC_data(
  ec_data = ec_data,
  blh_data = blh_data,
  required_vars = c("wind_speed", "v_var", "u.", "wind_dir", "blh", "L"),
  datetime_col = "datetime",
  filter_start_date = start_date,
  filter_end_date = end_date
)

# Rename u. to ustar for MATLAB compatibility
names(ec_clean)[names(ec_clean) == "u."] <- "ustar"

# Save prepared data for MATLAB
matlab_input_csv <- file.path(output_dir, "ec_data_prepared.csv")
write.csv(ec_clean, matlab_input_csv, row.names = FALSE)

message("Prepared ", nrow(ec_clean), " observations for MATLAB processing")
```


# Step 4: Calculate Footprint in MATLAB
```{r calc-footprint-matlab}
# Site parameters
zm <- 2.6
radius_vector <- c(50, 60, 70, 80, 90)

# Choose FFP function
# Options: "calc_footprint_FFP_climatology" or "calc_footprint_FFP_climatology_parallel"
ffp_function <- "calc_footprint_FFP_climatology_parallel"

# Create MATLAB command
matlab_cmd <- sprintf(
  "addpath('%s'); calc_footprint_matlab('%s', '%s', %.6f, %.6f, %.1f, [%s], '%s');",
  matlab_ffp_path,
  matlab_input_csv,
  output_dir,
  site_coords$long,
  site_coords$lat,
  zm,
  paste(radius_vector, collapse = ", "),
  ffp_function
)

# Run MATLAB
message("Running MATLAB footprint calculation with ", ffp_function, "...")
run_matlab_code(matlab_cmd)
message("MATLAB calculation complete!")
```

# Step 5: Load MATLAB Output and Convert to Shapefiles
```{r load-matlab-output}
# Read MATLAB output CSVs
footprint_list <- list()

for (pct in radius_vector) {
  csv_file <- file.path(output_dir, sprintf("footprint_%dpct.csv", pct))
  
  if (file.exists(csv_file)) {
    df <- read.csv(csv_file)
    
     # Remove rows with NaN/NA values
    df <- df[complete.cases(df), ]
    # Convert to shapefile using existing function
    footprint_sf <- long_lat_to_shapefile(
      df,
      project_shapefile = TRUE,
      project_crs_epsg = 3006
    )
    
    # Add contour percentage attribute
    footprint_sf$contour_pct <- pct
    
    footprint_list[[paste0(pct, "pct")]] <- footprint_sf
  } else {
    warning("MATLAB output file not found: ", csv_file)
  }
}

# Reverse order (largest first)
footprint_contours <- rev(footprint_list)

message("Loaded ", length(footprint_contours), " footprint contours from MATLAB output")
names(footprint_contours)
```

# Step 6: Visualize
```{r plot, fig.width=8, fig.height=7}
plot_footprint_with_site(
  footprint_sf = footprint_contours,
  site_coords = site_coords,
  title = "Footprint Climatology (MATLAB) - Halmyran",
  plot_crs_epsg = 3006,
  show_contour_labels = TRUE
)
```

# Step 7: Save Shapefiles (Optional)
```{r save-shapefiles, eval=FALSE}
# Define final output directory
final_output_dir <- "C:/my_output/footprints"

# Manually save using sf::st_write since we already have the sf objects
dir.create(final_output_dir, recursive = TRUE, showWarnings = FALSE)

# Save individual files
for (name in names(footprint_contours)) {
  file_path <- file.path(final_output_dir, paste0("footprint_", name, ".shp"))
  sf::st_write(footprint_contours[[name]], file_path, append = FALSE, quiet = TRUE)
}

# Save combined file
combined <- do.call(rbind, footprint_contours)
combined_path <- file.path(final_output_dir, "footprint_combined.shp")
sf::st_write(combined, combined_path, append = FALSE, quiet = TRUE)

message("Shapefiles saved to: ", final_output_dir)
```

# Comparison with R Implementation

The MATLAB implementation is significantly faster for large datasets (>1000 observations) but requires MATLAB to be installed. The R implementation (`calc_footprint_FFP_climatology.R`) is slower but doesn't require additional software.

**Choose MATLAB when:**
- Processing >1000 observations
- Speed is critical
- MATLAB is already available

**Choose R when:**
- Smaller datasets (<500 observations)
- Pure R workflow preferred
- MATLAB not available

# References

Kljun, N., Calanca, P., Rotach, M.W., Schmid, H.P., 2015. A simple two-dimensional parameterisation for Flux Footprint Prediction (FFP). Geoscientific Model Development 8, 3695–3713. https://doi.org/10.5194/gmd-8-3695-2015
```{r session-info}
sessionInfo()
```
