---
title: "Flux Footprint Analysis Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flux Footprint Analysis Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = TRUE
)
```

# Introduction

This vignette demonstrates flux footprint calculation and visualization using the Kljun et al. (2015) model with May 2020 data from the Halmyran peatland site in northern Sweden.

## Prerequisites

Download the FFP model from:
[http://footprint.kljun.net/](http://footprint.kljun.net/)
```{r eval=FALSE}
r_ffp_path <- system.file("FFP_R", package = "FFPUtilities")
ffp_climatology_function_path <- file.path(r_ffp_path, "calc_footprint_FFP_climatology.R")

source(ffp_climatology_function_path)
```

## Setup
```{r setup}
library(FFPUtilities)
library(ggplot2)
library(sf)
library(lubridate)


start_date = "2020-05-15"
end_date = "2020-05-16"
```

# Step 1: Load Eddy Covariance Data
```{r load-data}
ec_file <- system.file("extdata", "ec_data.csv", package = "FFPUtilities")
ec_data <- read.csv(ec_file)

ec_data$datetime <- parse_date_time(
  ec_data$datetime, 
  orders = c("ymd HMS", "ymd"), 
  tz = "UTC"
)


site_file <- system.file("extdata", "site_coords.csv", package = "FFPUtilities")
site_coords <- read.csv(site_file)

head(ec_data)
site_coords
```

# Step 2: Extract BLH from GRIB
```{r extract-blh}
grib_file <- system.file("extdata", "degfert_blh.grib", package = "FFPUtilities")

blh_data <- extract_blh_from_grib(
  grib_file = grib_file,
  site_long = site_coords$long,
  site_lat = site_coords$lat,
  filter_start_date = start_date,
  filter_end_date = end_date
)

head(blh_data)
```

# Step 3: Prepare Data
```{r prepare-data}
ec_clean <- prepare_EC_data(
  ec_data = ec_data,
  blh_data = blh_data,
  required_vars = c("wind_speed", "v_var", "u.", "wind_dir", "blh", "L"),
  datetime_col = "datetime",
  filter_start_date = start_date,
  filter_end_date = end_date
)

head(ec_clean)
```

# Step 4: Extract Parameters
```{r extract-params}
params <- extract_FFP_params(
  ec_data = ec_clean,
  row_index = 1:nrow(ec_clean),
  zm = 2.6,
  ustar_col = "u."
)

str(params)
```

# Step 5: Calculate Footprint
```{r calc-footprint, eval=FALSE}
radius_vector <- c(50, 60, 70, 80, 90)

FFP <- calc_footprint_FFP_climatology(
  zm = params$zm,
  z0 = params$z0,
  umean = params$umean,
  h = params$h,
  ol = params$ol,
  sigmav = params$sigmav,
  ustar = params$ustar,
  wind_dir = params$wind_dir,
  dx = 0.3,
  r = radius_vector,
  crop = 1,
  fig = 0
)
```
```{r simulate-ffp, echo=FALSE}
# set.seed(123)
# n_points <- 100
# 
# FFP <- list(
#   xr = lapply(1:5, function(i) {
#     angle <- seq(0, 2*pi, length.out = n_points)
#     radius <- (100 - i*10) * (1 + 0.3 * cos(3*angle))
#     radius * cos(angle + pi/6)
#   }),
#   yr = lapply(1:5, function(i) {
#     angle <- seq(0, 2*pi, length.out = n_points)
#     radius <- (100 - i*10) * (1 + 0.3 * cos(3*angle))
#     radius * sin(angle + pi/6) * 0.6
#   }),
#   x_2d = matrix(seq(-150, 150, length.out = 60), nrow = 1),
#   y_2d = matrix(seq(-150, 150, length.out = 60), ncol = 1),
#   fclim_2d = matrix(exp(-((seq(-150, 150, length.out = 60))^2)/5000), 
#                     nrow = 60, ncol = 60)
# )
# radius_vector <- c(50, 60, 70, 80, 90)
```

# Step 6: Convert to Shapefiles
```{r create-shapefiles}
footprint_contours <- FFP_to_shapefiles(
  FFP = FFP,
  site_long = site_coords$long,
  site_lat = site_coords$lat,
  radius_vector = radius_vector,
  project_shapefile = TRUE,
  project_crs_epsg = 3006,
  save_option = "both",
  output_dir = "C:/Users/kono0001/OneDrive - Sveriges lantbruksuniversitet/OTHER/personal/helping_others/Betty/Skogaryd/Scripts/Footprints"
)

names(footprint_contours)
```

# Step 7: Visualize
```{r plot, fig.width=8, fig.height=7}
plot_footprint_with_site(
  footprint_sf = footprint_contours,
  site_coords = site_coords,
  title = paste("Halmyran -", params$datetime),
  plot_crs_epsg = 3006,
  show_contour_labels = TRUE
)
```

# References

Kljun, N., Calanca, P., Rotach, M.W., Schmid, H.P., 2015. A simple two-dimensional parameterisation for Flux Footprint Prediction (FFP). Geoscientific Model Development 8, 3695–3713. https://doi.org/10.5194/gmd-8-3695-2015
```{r session-info}
sessionInfo()
```
